---
title: "Assurance for an Inteirm Look - DTE Example"
output: html_document
runtime: shiny
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rpact)
library(survival)
```

```{r functions, echo = F}

sim_dte <- function(n_c, n_t, lambda_c, delay_time, post_delay_HR, dist = "Exponential", gamma_c = NULL){

  #Simulate control data
  u <- runif(n_c)

  if (dist == "Exponential"){
    control_times <- -log(u)/lambda_c
  }

  if (dist == "Weibull"){
    control_times <- (1/lambda_c)*(-log(u))^(1/gamma_c)
  }

  #Simulate treatment data
  u <- runif(n_t)
  if (dist == "Exponential"){
    CP <- exp(-lambda_c*delay_time)
    treatment_times <- ifelse(u>CP,
                              -log(u)/lambda_c,
                              (1/(post_delay_HR*lambda_c))*(post_delay_HR*lambda_c*delay_time-log(u)-lambda_c*delay_time))
  }

  if (dist == "Weibull"){
    CP <- exp(-(lambda_c*delay_time)^gamma_c)
    lambda_e <- lambda_c*post_delay_HR^(1/gamma_c)
    treatment_times <- ifelse(u > CP,
                              (1/lambda_c)*(-log(u))^(1/gamma_c),
                              (1/lambda_e)*(-log(u)+(lambda_e*delay_time)^gamma_c-(lambda_c*delay_time)^gamma_c)^(1/gamma_c))
  }

  #Combine the two groups
  data <- data.frame(time = c(control_times, treatment_times),
                     group = c(rep("Control", n_c), rep("Treatment", n_t)))

  return(data)

}

add_recruitment_time <- function(data, rec_method,
                                 rec_period=NULL, rec_power=NULL, rec_rate=NULL, rec_duration=NULL){

  n_patients <- nrow(data)

  if (rec_method=="power"){

    data$rec_time <- rec_period * stats::runif(n_patients)^(1/rec_power)

  }

  if (rec_method == "PWC") {
    # Parse recruitment rate and duration inputs
    rec_rate <- as.numeric(unlist(strsplit(rec_rate, ",")))
    rec_duration <- as.numeric(unlist(strsplit(rec_duration, ",")))

    # Ensure valid inputs
    if (any(rec_rate < 0)) stop("rec_rate should be non-negative")
    if (length(rec_rate) != length(rec_duration)) stop("Lengths of rec_rate and rec_duration should match")

    n_periods <- length(rec_duration)

    if (length(rec_rate) == 1) { # Simple case with only one rate
      rec <- cumsum(stats::rexp(n = n_patients, rate = rec_rate))
    } else { # Piecewise recruitment
      # Create a data frame for the piecewise periods
      df <- data.frame(
        rate = rec_rate,
        duration = rec_duration,
        period = 1:n_periods,
        finish = cumsum(rec_duration),
        lambda = rec_duration * rec_rate,
        origin = c(0, cumsum(rec_duration)[-n_periods])
      )

      # Generate the number of recruits in each period using Poisson distribution
      df$N <- sapply(df$lambda, function(x) stats::rpois(n = 1, lambda = x))

      # Check if any recruits were generated
      if (sum(df$N) == 0) {
        if (df$rate[n_periods] == 0) stop("Please specify positive rec_rate for the last period; otherwise, enrollment cannot finish.")
        rec <- cumsum(stats::rexp(n = n_patients, rate = df$rate[n_periods])) + df$finish[n_periods]
      } else {
        # Generate recruitment times for each period
        rec <- unlist(apply(df, 1, function(x) {
          sort(stats::runif(n = x[["N"]], min = x[["origin"]], max = x[["finish"]]))
        }))

        # Check if we have enough recruits
        if (length(rec) >= n_patients) {
          rec <- rec[1:n_patients]
        } else {
          # Ensure enrollment completion if needed
          if (df$rate[n_periods] == 0) stop("Please specify positive rec_rate for the last period; otherwise, enrollment cannot finish.")

          # Generate additional recruitment times if needed
          rec <- c(rec, cumsum(stats::rexp(n_patients - length(rec), rate = df$rate[n_periods])) + df$finish[n_periods])
        }
      }
    }

    # Assign recruitment times to the data frame
    data$rec_time <- rec
  }

  data$pseudo_time <- data$time + data$rec_time

  return(data)

}

cens_data <- function(data, cens_method = "Time", cens_events = NULL, cens_time = NULL){

  if (cens_method=="Events"){
    data <- data[order(data$pseudo_time),]
    cens_time <- data$pseudo_time[cens_events]
  }

  data$status <- data$pseudo_time <= cens_time
  data$status <- data$status * 1
  data$enrolled <- data$rec_time < cens_time
  data <- data[data$enrolled, ]
  data$survival_time <- ifelse(data$pseudo_time > cens_time,
                               cens_time - data$rec_time,
                               data$time)

  return(list(data = data,
              cens_events = cens_events,
              cens_time = cens_time,
              sample_size = nrow(data)))
}
```

### Setup

We are planning a trial in which we anticipate will be subject to a delayed treatment effect. We have elicited distributions for key parameters, here: the length of delay, and the post-delay hazard ratio. We will use these distributions to plan an adaptive clinical trial. The parameters are:

```{r parameters}
lambda_c <- 0.08
P_S <- 1
P_DTE <- 1
delay_time_SHELF <- SHELF::fitdist(c(2.99, 3, 3.01), probs = c(0.25, 0.5, 0.75), lower = 0, upper = 10)
post_delay_HR_SHELF <- SHELF::fitdist(c(0.549, 0.55, 0.551), probs = c(0.25, 0.5, 0.75), lower = 0, upper = 1.5)
post_delay_HR_dist <- "gamma"
delay_time_dist <- "gamma"

```

and the parameters relating to the trial are:

```{r}
n_c <- 300
n_t <- 300
rec_method <- "power"
rec_period <- 12
rec_power <- 1
cens_method <- "Events"
cens_events <- 450
```

Here, we set the information fractions we will look at and the number of simulations we will perform:

```{r sims}
IF <- seq(0.1, 1, by=0.1)
NSims <- 1e1
```

```{r inits, echo=F}

library(doParallel)
library(foreach)
library(SHELF)
library(survival)

# Register parallel backend
n_cores <- parallel::detectCores() - 1
cl <- makeCluster(n_cores)
registerDoParallel(cl)

# Allocate storage
delay_vec <- numeric(NSims)
post_delay_HR_vec <- numeric(NSims)
z_scores_mat <- matrix(NA, nrow = length(IF), ncol = NSims)
duration_mat <- matrix(NA, nrow = length(IF), ncol = NSims)
ss_mat <- matrix(NA, nrow = length(IF), ncol = NSims)
final_duration_mat <- matrix(NA, nrow = length(IF), ncol = NSims)
final_ss_mat <- matrix(NA, nrow = length(IF), ncol = NSims)

# Run simulation in parallel
results <- foreach(j = 1:NSims, .packages = c("SHELF", "survival")) %dopar% {

  if (runif(1) > P_S) {
    delay_time <- 0
    post_delay_HR <- 1
  } else {
    if (runif(1) > P_DTE) {
      delay_time <- 0
      post_delay_HR <- SHELF::sampleFit(post_delay_HR_SHELF, n = 1)[, post_delay_HR_dist]
    } else {
      delay_time <- SHELF::sampleFit(delay_time_SHELF, n = 1)[, delay_time_dist]
      post_delay_HR <- SHELF::sampleFit(post_delay_HR_SHELF, n = 1)[, post_delay_HR_dist]
    }
  }

  data <- sim_dte(n_c, n_t, lambda_c, delay_time, post_delay_HR, dist = "Exponential")

  if (rec_method == "power") {
    data <- add_recruitment_time(data, rec_method, rec_period, rec_power)
  } else if (rec_method == "PWC") {
    data <- add_recruitment_time(data, rec_method, rec_rate, rec_duration)
  }

  final_data <- data
  z_scores <- duration <- ss <- numeric(length(IF))

  for (k in seq_along(IF)) {
    data_after_cens <- cens_data(data = final_data, cens_method = cens_method,
                                 cens_events = cens_events * IF[k])
    data_k <- data_after_cens$data

    coxmodel <- coxph(Surv(survival_time, status) ~ group, data = data_k)
    z_scores[k] <- -coef(summary(coxmodel))[, 4]
    duration[k] <- data_after_cens$cens_time
    ss[k] <- data_after_cens$sample_size
  }

  list(
    delay_time = delay_time,
    post_delay_HR = post_delay_HR,
    z_scores = z_scores,
    duration = duration,
    ss = ss
  )
}

# Collect results
for (j in 1:NSims) {
  delay_vec[j] <- results[[j]]$delay_time
  post_delay_HR_vec[j] <- results[[j]]$post_delay_HR
  z_scores_mat[, j] <- results[[j]]$z_scores
  duration_mat[, j] <- results[[j]]$duration
  ss_mat[, j] <- results[[j]]$ss
}

stopCluster(cl)

```

### No Interim Analysis

First, we calculate the assurance (in the case we do not perform any interim analyses).

```{r, echo = F}


design <- getDesignGroupSequential(typeOfDesign = "asUser",
                                   informationRates = 1,
                                   userAlphaSpending = 0.025)


final_outcome <- ifelse(z_scores_mat[length(IF),]>design$criticalValues, 1, 0)

mean(final_outcome)

```

We plot the sampled Length of Delay vs post-delay HR. The points are coloured by the success of the trial (statistical significance at the final analysis).

```{r, echo = F}
outcome_labels <- ifelse(final_outcome == 1, "Successful at Final Analysis",
                         "Unsuccessful at Final Analysis")
outcomes <- c("Successful at Final Analysis", "Unsuccessful at Final Analysis")
outcome_colors <- setNames(c("blue", "red"), outcomes)
point_colors <- outcome_colors[outcome_labels]


plot(delay_vec, post_delay_HR_vec, col = point_colors, cex = 0.5,
     xlab = "Length of Delay", ylab = "Post-Delay HR")
legend("topright",
       legend = outcomes,
       col = c("blue", "red"),
       pch = 1,
       cex = 0.8)


```

### One Interim Analysis

Now, we assume we perform ONE interim analysis for each of the simulated trials. There are four different outcomes that each trial can have:

* Stop for Efficacy
* Stop for Futility
* Successful at the final analysis (implicitly implied that we continued at the IA)
* Unsuccessful at the final analysis (implicitly implied that we continued at the IA)

```{r, echo = F}

Trial_Outcome_mat <- data.frame(matrix(NA, ncol = NSims, nrow = length(IF)-1))


for (k in 1:(length(IF)-1)){
  design <- getDesignGroupSequential(typeOfDesign = "asUser",
                                     informationRates = c(IF[k], 1),
                                     userAlphaSpending = c(0.0125, 0.025),
                                     typeBetaSpending = "bsUser",
                                     userBetaSpending = c(0.05, 0.1))
 
  for (i in 1:NSims){
    IA_ZScore <- z_scores_mat[k,i]
    Final_ZScore <- z_scores_mat[length(IF), i]
    Trial_Outcome_mat[k,i] <- ifelse(IA_ZScore > design$criticalValues[1], "Stop for Efficacy",
                                     ifelse(IA_ZScore < design$futilityBounds, "Stop for Futility",
                                            ifelse(Final_ZScore > design$criticalValues[2],
                                                   "Successful at Final Analysis",
                                                   "Unsuccessful at Final Analysis")))
    
    final_duration_mat[k,i] <- ifelse(IA_ZScore > design$criticalValues[1], duration_mat[k,i],
                                     ifelse(IA_ZScore < design$futilityBounds, duration_mat[k,i],
                                            duration_mat[length(IF),i]))
    
    final_ss_mat[k,i] <- ifelse(IA_ZScore > design$criticalValues[1], ss_mat[k,i],
                                     ifelse(IA_ZScore < design$futilityBounds, ss_mat[k,i],
                                            ss_mat[length(IF),i]))
    
    
    
  }
 
 
}

final_duration_mat[length(IF), ] <- duration_mat[length(IF), ]
final_ss_mat[length(IF), ] <- ss_mat[length(IF), ]



```

Again, we plot the same sampled parameters and colour the points the outcome each trial had. By changing the Information Fraction, we change the decisions of a trial. 

```{r, echo = F}

outcomes <- c("Successful at Final Analysis", "Stop for Efficacy", "Stop for Futility", "Unsuccessful at Final Analysis")
outcome_colors <- setNames(c("green", "green", "red", "red"), outcomes)


sliderInput("chosen_IF", "Information Fraction", min = IF[1],
            max = IF[length(IF)-1], value = IF[1], step = IF[2] - IF[1])


renderPlot({
  
    point_colors <- outcome_colors[as.character(Trial_Outcome_mat[which.min(abs(IF - input$chosen_IF))
, ])]
    
    #print(point_colors)
 
  plot(delay_vec, post_delay_HR_vec,
       col = point_colors,
       cex = 0.5,
       xlab = "Length of Delay",
       ylab = "Post-Delay HR",
       main = paste0("Outcomes when we perform IA at: ", input$chosen_IF))
 
  # Add legend
  legend("topright",                               # Position of the legend
         legend = outcomes,                        # Text labels
         col = outcome_colors[outcomes],           # Corresponding colors
         pch = 1,                                  # Point type to match plot
         cex = 0.8)                                 # Size of legend text
})
  
  




```

We also show this in table format
```{r, echo = F}

IF_outcomes_mat <- data.frame(matrix(NA, ncol = length(IF), nrow = length(outcomes)))
colnames(IF_outcomes_mat) <- IF
rownames(IF_outcomes_mat) <- outcomes

for (i in 1:length(IF)){
  for (k in 1:length(outcomes)){
    IF_outcomes_mat[k,i] <- mean(Trial_Outcome_mat[i,]==outcomes[k])
  }
}

for (k in 1:length(outcomes)){
  IF_outcomes_mat[k,length(IF)] <- mean(ifelse(final_outcome==1,
                                               "Successful at Final Analysis",
                                               "Unsuccessful at Final Analysis")==outcomes[k])
                                                 
                                                 
}




```

And a stacked barchart.

```{r, echo = F}
# Adjust margins: c(bottom, left, top, right)
density_vals <- c(NA, 50, 50, NA)  
angle_vals <- c(0, 0, 0, 0) 
par(mar = c(5, 4, 4, 15))  # Increase right margin

barplot(as.matrix(IF_outcomes_mat),
        beside = FALSE,
        col = outcome_colors[outcomes],
        density = density_vals,
        angle = angle_vals,
        xlab = "Information Fraction",
        ylab = "Proportion",
        main = "Outcomes at each IA")

legend("topright",
       legend = outcomes,
       fill = outcome_colors[outcomes],
       cex = 0.6,
       density = density_vals,
        angle = angle_vals,
       xpd = TRUE,  # Allows legend to plot outside plot area
       inset = c(-0.5, 0))  # Adjust based on margin size
```


As we have access to the final results, we can also calculate whether a decision made at the interim analysis was correct (or not). As such, we have six outcomes for our trial:

* Correctly stopped for futility 
* Incorrectly stopped for futility
* Correctly stopped for efficacy 
* Incorrectly stopped for efficacy
* Successful at the final analysis
* Unsuccessful at the final analysis


Same plot, with the same colours as before, but we use different shapes to show the (in)correct decisions.

```{r, echo = F}


Trial_Outcome_decisions_mat <- Trial_Outcome_mat

for (j in 1:nrow(Trial_Outcome_mat)){
  for (k in 1:ncol(Trial_Outcome_mat)){
    if (Trial_Outcome_mat[j,k]=="Stop for Efficacy"){
      Trial_Outcome_decisions_mat[j,k] <- ifelse(final_outcome[k]==1,
                                                 "Correctly stopped for Efficacy",
                                                 "Incorrectly stopped for Efficacy")
    }
    if (Trial_Outcome_mat[j,k]=="Stop for Futility"){
      Trial_Outcome_decisions_mat[j,k] <- ifelse(final_outcome[k]==0,
                                                 "Correctly stopped for Futility",
                                                 "Incorrectly stopped for Futility")
    }
  }
}

```


```{r, echo = F}


outcomes_decisions <- c("Successful at Final Analysis", "Correctly stopped for Efficacy",
                        "Incorrectly stopped for Efficacy", "Incorrectly stopped for Futility",
                        "Correctly stopped for Futility", "Unsuccessful at Final Analysis")

outcome_decisions_colors <- setNames(c("green", "green", "green", "red", "red", "red"), outcomes_decisions)


outcome_decisions_shapes <- setNames(c(1, 1, 0, 2, 0, 2), outcomes_decisions)


sliderInput("chosen_IF_decisions", "Information Fraction", min = IF[1],
            max = IF[length(IF)-1], value = IF[1], step = IF[2] - IF[1])


renderPlot({
  
    point_colors_decisions <- outcome_decisions_colors[as.character(Trial_Outcome_decisions_mat[which.min(abs(IF - input$chosen_IF_decisions))
, ])]
    
    point_shapes_decisions <- outcome_decisions_shapes[as.character(Trial_Outcome_decisions_mat[which.min(abs(IF - input$chosen_IF_decisions))
, ])]
    
 
  plot(delay_vec, post_delay_HR_vec,
       col = point_colors_decisions,
       cex = 0.5,
       pch = point_shapes_decisions,
       xlab = "Length of Delay",
       ylab = "Post-Delay HR",
       main = paste0("Outcomes when we perform IA at: ", input$chosen_IF_decisions))
 
  # Add legend
  legend("topright",                               # Position of the legend
         legend = outcomes_decisions,                        # Text labels
         col = outcome_decisions_colors[outcomes_decisions],           # Corresponding colors
         pch = c(1, 1, 0, 2, 0, 2),                                  # Point type to match plot
         cex = 0.8)                                 # Size of legend text
})
  
  

```


Table format:

```{r, echo = F}

IF_outcomes_mat <- data.frame(matrix(NA, ncol = length(IF), nrow = length(outcomes_decisions)))
colnames(IF_outcomes_mat) <- IF
rownames(IF_outcomes_mat) <- outcomes_decisions

for (i in 1:length(IF)){
  for (k in 1:length(outcomes_decisions)){
    IF_outcomes_mat[k,i] <- mean(Trial_Outcome_decisions_mat[i,]==outcomes_decisions[k])
  }
}

for (k in 1:length(outcomes_decisions)){
  IF_outcomes_mat[k,length(IF)] <- mean(ifelse(final_outcome==1,
                                               "Successful at Final Analysis",
                                               "Unsuccessful at Final Analysis")==outcomes_decisions[k])
                                                 
                                                 
}

kable(IF_outcomes_mat, digits = 2)


```

And the stacked barchart. We use the same colours as the first barchart, but each colour is broken down into correct and incorrect decision.

```{r, echo = F}

png("DTE_Interim_Example.png", units="in", width=10, height=6, res=700)
density_vals <- c(NA, 20, 50, 50, 20, NA)  
angle_vals <- c(0, 45, -45, -45, 45, 0)     

par(mar = c(5, 4, 4, 20))  # Increase right margin

# Get bar heights to determine where to draw lines
bar_heights <- apply(IF_outcomes_mat[1:3, ], 2, sum)


bar_positions <- barplot(as.matrix(IF_outcomes_mat),
                         beside = FALSE,
                         width = 1,  # Optional, but explicit
                         col = outcome_decisions_colors[outcomes_decisions],
                         density = density_vals,
                         angle = angle_vals,
                         xlab = "Information Fraction",
                         ylab = "Proportion",
                         main = "Proportion of Trial Outcomes at Different Information Fractions")



legend("topright",
       legend = c(rev(outcomes_decisions), "Assurance"),
       fill = c(rev(outcome_decisions_colors[outcomes_decisions]), NA),
       border = c(rep("black", length(outcomes_decisions)), NA),
       density = c(rev(density_vals), NA),
       angle = c(rev(angle_vals), NA),
       col = c(rep(NA, length(outcomes_decisions)), "black"),
       cex = 0.6,
       xpd = TRUE,
       inset = c(-0.4, 0),
       bty = "n")



# Add black separator lines between top and bottom outcome segments
segments(x0 = bar_positions - 0.5,  # left edge of each bar
         x1 = bar_positions + 0.5,  # right edge
         y0 = bar_heights,          # y position (cumulative height of top 3)
         y1 = bar_heights,
         col = "black",
         lwd = 5)

# Add text labels just below the separator line
text(x = bar_positions,
     y = bar_heights,  # Adjust this for spacing
     labels = round(bar_heights, 2),
     cex = 1,                # text size
     pos = 1)                  # below the specified y (1 = below)

dev.off()


```

```{r}

IF_outcomes_mat <- rbind(IF_outcomes_mat, rowMeans(final_duration_mat), rowMeans(final_ss_mat))

rownames(IF_outcomes_mat) <- c(outcomes_decisions, "Expected Duration", "Expected Sample Size")

kable(IF_outcomes_mat, digits = 2)
```

