---
title: "Likelihood"
header-includes:
   - \usepackage{bm}
output: html_document
date: '2022-07-22'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(stats4)
```

## R Markdown

We have some data $\textbf{x}= (x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m})$. We are tasked with finding the most likely values for the parameters $\lambda_c$, $\lambda_t$ and $T$ if we assume they come from the following survival functions

$$S_c(t) = \text{exp}\{-\lambda_ct\}$$

$$S_t(t) = \text{exp}\{-\lambda_ct\}\mathbb{1}_{t\leq T}+\exp\{-\lambda_cT-\lambda_t(t-T)\}\mathbb{1}_{t>T}$$

We have the relationship

$$f(t) = \frac{d}{dt}[1-S(t)]$$

Therefore, the above survival probabilities can be manipulated to give the following densities

$$f_c(t) = \lambda_c\text{exp}\{-\lambda_ct\}$$
$$f_t(t) = \lambda_c\text{exp}\{-\lambda_ct\}\mathbb{1}_{t\leq T}+\lambda_t\exp\{-\lambda_cT-\lambda_t(t-T)\}\mathbb{1}_{t>T}$$

Therefore, the joint density is


\begin{align}
p(\textbf{x}|\lambda_c, \lambda_t,T) &= p(x_{c,1},\ldots,x_{c,n}|\lambda_c) \times p(x_{t,1},\ldots,x_{t,m}|\lambda_c,\lambda_t, T) \\
&= p(x_{c,1},\ldots,x_{c,n}|\lambda_c) \times p(x_{t,1},\ldots,x_{t,j}|\lambda_c) \times p(x_{t,j+1},\ldots,x_{t,m}|\lambda_c,\lambda_t, T),
\end{align}

where $x_{t,j}$ is the largest observed $x$ in the treatment group before T.

The joint likelihood is 

\begin{align}
\mathcal{L}(\lambda_c,\lambda_t, T|x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m}) &= \prod^n_{i=1} \lambda_c\text{exp}\{-\lambda_cx_{c,i}\} \times \prod^j_{k=1} \lambda_c\text{exp}\{-\lambda_cx_{t,k}\} \times \prod^m_{p=j+1} \lambda_t\text{exp}\{-\lambda_t(x_{t,p}-T)-\lambda_cT\} \\
&= \lambda_c^n\text{exp}(-\lambda_c \sum^n_{i=1}x_{c,i}) \times \lambda_c^j\text{exp}(-\lambda_c \sum^j_{k=1}x_{t,k}) \times \lambda_t^{m-j}\text{exp}\{(m-j)(\lambda_t T-\lambda_c T)-\lambda_t \sum^m_{p=j+1}x_{t,p}\}
\end{align}

We find the log-likelihood

\begin{align}
\text{log} \mathcal{L}(\lambda_c,\lambda_t, T|x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m}) &= n \text{log}\lambda_c-\lambda_c\sum^n_{i=1}x_{c,i}+j\text{log}\lambda_c-\lambda_c\sum^j_{k=1}x_{t,k}+(m-j)\text{log}\lambda_t+(m-j)(\lambda_t T-\lambda_c T)-\lambda_t\sum^m_{p=j+1}x_{t,p} \\
&= (n+j)\text{log}\lambda_c - \lambda_c\big(\sum^n_{i=1}x_{c,i}+\sum^j_{k=1}x_{t,k}\big) +(m-j)(\text{log}\lambda_t+\lambda_t T - \lambda_c T)-\lambda_t \sum^m_{p=j+1}x_{t,p}
\end{align}

We can ensure that this is the correct log-likelihood with the following code

```{r}
#Setting up the correct parameters - we would not know this in practice

lambda1 <- 0.04
lambda2 <- 0.08
bigT <- 6

#Sample sizes in each group
n1 <- 100
n2 <- 100

#Simulating the control and treatment data - again, we would not normally know the underlying structure, but for illustration this is okay
#Control
controlData <- data.frame(time = rexp(n1, lambda2))
#Treatment
CP <- exp(-(lambda2*bigT))[[1]]
u <- runif(n2)
suppressWarnings(treatmentData <- ifelse(u>CP, -(1/lambda2)*log(u), (1/lambda1)*(-log(u)-bigT*(lambda2-lambda1))))

#Using the data to find the parameters we have defined
n <- length(controlData$time)
m <- length(treatmentData)
treatmentData <- sort(treatmentData)

#Coding the log-likelihood function we derived
nll <- function(bigTEst, lambda2Est, lambda1Est){
  j <- sum(treatmentData < bigTEst)
  loglikelihood <- (n+j)*log(lambda2Est)-lambda2Est*(sum(controlData$time)+sum(treatmentData[1:j]))+
  (m-j)*(log(lambda1Est)+lambda1Est*bigTEst-lambda2Est*bigTEst) - lambda1Est*sum(treatmentData[(j+1):m])
  #Returning the negative ll as this is what the mle() function expects
 return(-loglikelihood)
}

#Finding the maximum likelihood estimators for T, lambda2 and lambda1
est <- stats4::mle(minuslogl = nll, start = list(bigTEst=6, lambda2Est = 0.01, lambda1Est = 0.01), lower=  list(bigTEst=0.5, lambda2Est = 0.0001, lambda1Est = 0.0001), upper = list(bigTEst = 60))

#Showing the output of the mle() function
#Seems pretty good!
est

```

```{stan output.var="ia1", eval=F}

functions {
  real expmodel_log(vector control, vector treatment, real lambdac, real lambdat, real bigT) {
    real loglik;
    int j;
    int n;
    int m;
    j <- sum(treatment<bigT);
    n <- length(control)
    m <- length(treatment)
    loglik <- (n+j)*log(lambdac)-lambdac*(sum(control)+sum(treatment[1:j]))+
  (m-j)*(log(lambdat)+lambdat*bigT-lambdac*bigT) - lambdat*sum(treatment[(j+1):m])
    return loglik;
  }
}

data {
  int LENGTH;
  vector[LENGTH] Y;
}

parameters {
  real <lower=0> lambdac;
  real <lower=0> lambdat;
  real <lower=0> bigT;
}

```

https://my.vanderbilt.edu/jeffannis/files/2016/06/AnnisMillerPalmeri2016.pdf




