---
title: "LikelihoodWeib"
header-includes:
   - \usepackage{bm}
output: html_document
date: '2022-07-22'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(stats4)
library(rstan)
library(ggmcmc)
```

## R Markdown

We have some data $\textbf{x}= (x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m})$. We are tasked with finding the most likely values for the parameters $\lambda_1$, $\gamma_1$, $\lambda_2$, $\gamma_2$ and $T$ if we assume they come from the following survival functions

$$S_c(t) = \text{exp}\{-(\lambda_2t)^{\gamma_2}\}$$

$$S_t(t) = \text{exp}\{-(\lambda_2t)^{\gamma_2}\}\mathbb{1}_{t\leq T}+\exp\{-(\lambda_2T)^{\gamma_2}-\lambda_1^{\gamma_1}(t^{\gamma_1}-T^{\gamma_1})\}\mathbb{1}_{t>T}$$

We have the relationship

$$f(t) = \frac{d}{dt}[1-S(t)]$$

Therefore, the above survival probabilities can be manipulated to give the following densities

$$f_c(t) = \frac{\gamma_2(\lambda_2t)^{\gamma_2}\text{exp}\{-(\lambda_2t)^{\gamma_2}\}}{t}$$
$$f_t(t) = \frac{\gamma_2(\lambda_2t)^{\gamma_2}\text{exp}\{-(\lambda_2t)^{\gamma_2}\}}{t}\mathbb{1}_{t\leq T}+\gamma_1\lambda_1^{\gamma_1}t^{\gamma_1-1}\text{exp}\{-\lambda_1^{\gamma_1}(t^{\gamma_1}-T^{\gamma_1})-(T\lambda_2)^{\gamma_2}\}\mathbb{1}_{t>T}$$

Therefore, the joint density is


\begin{align}
p(\textbf{x}|\lambda_1, \gamma_1,\lambda_2,\gamma_2, T) &= p(x_{c,1},\ldots,x_{c,n}|\lambda_2, \gamma_2) \times p(x_{t,1},\ldots,x_{t,m}|\lambda_1, \gamma_1,\lambda_2,\gamma_2, T) \\
&= p(x_{c,1},\ldots,x_{c,n}|\lambda_2, \gamma_2) \times p(x_{t,1},\ldots,x_{t,j}|\lambda_2, \gamma_2) \times p(x_{t,j+1},\ldots,x_{t,m}|\lambda_1, \gamma_1,\lambda_2,\gamma_2, T),
\end{align}

where $x_{t,j}$ is the largest observed $x$ in the treatment group before T.

The joint likelihood is 

\begin{align}
\mathcal{L}(\lambda_1, \gamma_1,\lambda_2,\gamma_2, T|x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m}) &= \prod^n_{i=1} \frac{\gamma_2(\lambda_2x_{c,i})^{\gamma_2}\text{exp}\{-(\lambda_2x_{c,i})^{\gamma_2}\}}{x_{c,i}} \times \prod^j_{k=1}  \frac{\gamma_2(\lambda_2x_{t,k})^{\gamma_2}\text{exp}\{-(\lambda_2x_{t,k})^{\gamma_2}\}}{x_{t,k}} \times \prod^m_{p=j+1}
\gamma_1\lambda_1^{\gamma_1}x_{t,p}^{\gamma_1-1}\text{exp}\{-\lambda_1^{\gamma_1}(x_{t,p}^{\gamma_1}-T^{\gamma_1})-(T\lambda_2)^{\gamma_2}\} \\


&= \gamma_2^n\lambda_2^{n\gamma_2}\{\prod_{i=1}^nx_{c,i}\}^{\gamma_2}\text{exp}\{-\lambda_2^{\gamma_2}\sum^n_{i=1}(x_{c,i}^{\gamma_2})\}\prod^n_{i=1}\frac{1}{x_{c,i}} \times \gamma_2^j\lambda_2^{j\gamma_2}\{\prod_{k=1}^jx_{t,k}\}^{\gamma_2}\text{exp}\{-\lambda_2^{\gamma_2}\sum^j_{k=1}(x_{t,k}^{\gamma_2})\}\prod^j_{k=1}\frac{1}{x_{t,k}} \times \gamma_1^{m-j}\lambda_1^{\gamma_1(m-j)}\{\prod^m_{p=j+1}x_{t,p}\}^{\gamma_1-1}\text{exp}\{-\lambda_1^{\gamma_1}\sum^m_{p=j+1}(x_{t,p}^{\gamma_1})+(m-j)\big[(\lambda_1T)^{\gamma_1}-(\lambda_2T)^{\gamma_2}\big]\}
\end{align}

We find the log-likelihood

\begin{align}
\text{log} \mathcal{L}(\lambda_1, \gamma_1,\lambda_2,\gamma_2, T|x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m}) &= n\text{log}\gamma_2 + (n\gamma_2)\text{log}\lambda_2+\gamma_2\sum^n_{i=1}\text{log}x_{c,i}-\lambda_2^{\gamma_2}\sum^{n}_{i=1}x_{c,i}^{\gamma_2}+\sum^{n}_{i=1}\text{log}\frac{1}{x_{c,i}} + j\text{log}\gamma_2 + (j\gamma_2)\text{log}\lambda_2+\gamma_2\sum^j_{k=1}\text{log}x_{t,k}-\lambda_2^{\gamma_2}\sum^{j}_{k=1}x_{t,k}^{\gamma_2}+\sum^{j}_{k=1}\text{log}\frac{1}{x_{t,k}}+(m-j)\text{log}\gamma_1+\gamma_1(m-j)\text{log}\lambda_1+(\gamma_1-1)\sum^m_{p=j+1}\text{log}x_{t,p} - \lambda_1^{\gamma_1}\sum^m_{p=j+1}(x_{t,p}^{\gamma_1})+(m-j)\big[(\lambda_1T)^{\gamma_1}-(\lambda_2T)^{\gamma_2}\big]
\end{align}


We can ensure that this is the correct log-likelihood with the following code

```{r, eval=F}
#Setting up the correct parameters - we would not know this in practice

lambda1 <- 0.04
lambda2 <- 0.08
gamma2 <- 0.8
gamma1 <- 0.8
bigT <- 6

#Sample sizes in each group
n1 <- 1000
n2 <- 1000

#Simulating the control and treatment data - again, we would not normally know the underlying structure, but for illustration this is okay
#Control
controlData <- data.frame(time = rexp(n1, lambda2))
#Treatment
CP <- exp(-(lambda2*bigT))[[1]]
u <- runif(n2)
suppressWarnings(treatmentData <- ifelse(u>CP, -(1/lambda2)*log(u), (1/lambda1)*(-log(u)-bigT*(lambda2-lambda1))))

#Using the data to find the parameters we have defined
n <- length(controlData$time)
m <- length(treatmentData)
treatmentData <- sort(treatmentData)

#Coding the log-likelihood function we derived
nll <- function(bigTEst, lambda2Est, lambda1Est){
  j <- sum(treatmentData < bigTEst)
  loglikelihood <- (n+j)*log(lambda2Est)-lambda2Est*(sum(controlData$time)+sum(treatmentData[1:j]))+
  (m-j)*(log(lambda1Est)+lambda1Est*bigTEst-lambda2Est*bigTEst) - lambda1Est*sum(treatmentData[(j+1):m])
  #Returning the negative ll as this is what the mle() function expects
 return(-loglikelihood)
}

#Finding the maximum likelihood estimators for T, lambda2 and lambda1
est <- stats4::mle(minuslogl = nll, start = list(bigTEst=6, lambda2Est = 0.01, lambda1Est = 0.01), lower=  list(bigTEst=0.5, lambda2Est = 0.0001, lambda1Est = 0.0001), upper = list(bigTEst = 60))

#Showing the output of the mle() function
#Seems pretty good!
est

```

```{stan output.var="ia1", eval=F}

functions {
  real expmodel_log(vector Y, int controlLength, int treatmentLength, real lambdac, real lambdat, real bigT) {
    real loglik;
    int j;
    int n;
    int m;
    vector[controlLength] controlData;
    vector[treatmentLength] treatmentData;
    n = controlLength;
    m = treatmentLength;
    controlData = Y[1:n];
    treatmentData = Y[(n+1):(n+m)];
    j=0;
    for (i in 1:m){
      if (treatmentData[i]<bigT){
        j = j + 1;
      }
    }
   
    loglik = (n+j)*log(lambdac)-lambdac*(sum(controlData)+sum(treatmentData[1:j]))+
  (m-j)*(log(lambdat)+lambdat*bigT-lambdac*bigT) - lambdat*sum(treatmentData[(j+1):(m)]);
    return loglik;
  }
}

data {
  int controlLength;
  int treatmentLength;
  vector[controlLength+treatmentLength] Y;
}

parameters {
  real <lower=0> lambdac;
  real <lower=0> lambdat;
  real <lower=0> bigT;
}

model {
  lambdac ~ beta(1, 1);
  lambdat ~ beta(1, 1);
  bigT ~ normal(6, 1);
  Y ~ expmodel(controlLength, treatmentLength, lambdac, lambdat, bigT);
}

```


```{r, eval=F}

lambda1 <- 0.04
lambda2 <- 0.08
bigT <- 6

#Sample sizes in each group
n1 <- 500
n2 <- 500

#Simulating the control and treatment data - again, we would not normally know the underlying structure, but for illustration this is okay
#Control
controlData <-rexp(n1, lambda2)
#Treatment
CP <- exp(-(lambda2*bigT))[[1]]
u <- runif(n2)
suppressWarnings(treatmentData <- ifelse(u>CP, -(1/lambda2)*log(u), (1/lambda1)*(-log(u)-bigT*(lambda2-lambda1))))

n <- length(controlData)
m <- length(treatmentData)
treatmentData <- sort(treatmentData)

combinedData <- c(controlData, treatmentData)

 sampling_iterations <- 1e3
 fit <- sampling(ia1,
 data = list(Y = combinedData, controlLength=n, treatmentLength=m),
 chains = 1,
 iter = sampling_iterations,
 warmup = sampling_iterations/4,
 refresh=0)
 
 summary(fit)
 
 samples <- ggmcmc::ggs(fit)
 ggmcmc::ggs_pairs(samples)
 
 ggmcmc::ggs_caterpillar(samples, thick_ci = c(0.25,0.75))
 
 ggmcmc::ggs_traceplot(samples)
```


https://my.vanderbilt.edu/jeffannis/files/2016/06/AnnisMillerPalmeri2016.pdf


#Making sure the weibull likelihood is correct

```{r}


lambda1 <- 0.04
lambda2 <- 0.08
gamma2 <- 0.8
gamma1 <- 0.8
bigT <- 6

n1 <- 5
n2 <- 5

controldata <- rweibull(n1, gamma2, 1/lambda2)

CP <- exp(-(lambda2*bigT)^gamma2)[[1]]
u <- runif(n2)

suppressWarnings(treatmentdata <- ifelse(u>CP, (1/lambda2)*exp(1/gamma2*log(-log(u))), exp((1/gamma1)*log(1/(lambda1^gamma1)*(-log(u)-(lambda2*bigT)^gamma2+lambda1^gamma1*bigT*gamma1)))))

j <- sum(treatmentdata<bigT)


Like1 <- 1
Like2 <- 1
Like3 <- 1
for (i in 1:n1){
  Like1 <- Like1*((gamma2*(lambda2*controldata[i])^gamma2*exp(-(lambda2*controldata[i])^gamma2))/(controldata[i]))
}

for (i in 1:j){
  Like2 <- Like2*((gamma2*(lambda2*treatmentdata[i])^gamma2*exp(-(lambda2*treatmentdata[i])^gamma2))/(treatmentdata[i]))
}

for (i in (j+1):n2){
  Like3 <- Like3*(gamma1*lambda1^gamma1*treatmentdata[i]^(gamma1-1)*exp(-lambda1^gamma1*(treatmentdata[i]^gamma1-bigT^gamma1)-(bigT*lambda2)^gamma2))

}

Lik <- Like1*Like2*Like3


##Now time to see whether the simplification has worked
#Trying first part

mysum <- 0
for (i in 1:5){
  mysum <- mysum+ controldata[i]^gamma1
}

myprod <- 1
for (i in 1:5){
  myprod <- myprod *(1/controldata[i])
}

Siml1 <- gamma2^n1*lambda2^(n1*gamma2)*prod(controldata)^gamma2*exp(-lambda2^gamma2*mysum)*myprod

#Trying second part


mysum <- 0
for (i in 1:j){
  mysum <- mysum+ treatmentdata[i]^gamma1
}

myprod <- 1
for (i in 1:j){
  myprod <- myprod *(1/treatmentdata[i])
}

Siml2 <- gamma2^j*lambda2^(j*gamma2)*prod(treatmentdata[1:j])^gamma2*exp(-lambda2^gamma2*mysum)*myprod


#Trying third part

mysum <- 0

for (i in (j+1):n2){
  mysum <- mysum + treatmentdata[i]^gamma1
}


Siml3 <- lambda1^(n2-j)*lambda1^(gamma1*(n2-j))*prod(treatmentdata[(j+1):n2])^(gamma1-1)*exp(-lambda1^gamma1*mysum+)
#This bit is not quite finished



#Breaking it down

l <- 1
for (i in 1:5){
  l <-  l*(gamma2*(lambda2*controldata[i])^gamma2)
}

gamma2^5*lambda2^(5*gamma2)*prod(controldata)^gamma2







```





