---
title: "Bayesian analaysis of the DTE - at an interim analysis"
header-includes: \usepackage{bm}
date: 'September 2022'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(stats4)
library(rstan)
library(rjags)
library(ggmcmc)
```

We have some data $\textbf{x}= (x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m})$. We are tasked with finding the most likely values for the parameters $\lambda_1$, $\gamma_1$, $\lambda_2$, $\gamma_2$ and $T$ if we assume the data come from the following survival functions

$$S_c(t) = \text{exp}\{-(\lambda_2t)^{\gamma_2}\}$$

$$S_t(t) = \text{exp}\{-(\lambda_2t)^{\gamma_2}\}\mathbb{1}_{t\leq T}+\exp\{-(\lambda_2T)^{\gamma_2}-\lambda_1^{\gamma_1}(t^{\gamma_1}-T^{\gamma_1})\}\mathbb{1}_{t>T}$$

We have the relationship

$$f(t) = \frac{d}{dt}[1-S(t)]$$

Therefore, the above survival probabilities can be manipulated to give the following densities

$$f_c(t) = \frac{\gamma_2(\lambda_2t)^{\gamma_2}\text{exp}\{-(\lambda_2t)^{\gamma_2}\}}{t}$$ $$f_t(t) = \frac{\gamma_2(\lambda_2t)^{\gamma_2}\text{exp}\{-(\lambda_2t)^{\gamma_2}\}}{t}\mathbb{1}_{t\leq T}+\gamma_1\lambda_1^{\gamma_1}t^{\gamma_1-1}\text{exp}\{-\lambda_1^{\gamma_1}(t^{\gamma_1}-T^{\gamma_1})-(T\lambda_2)^{\gamma_2}\}\mathbb{1}_{t>T}$$

<!-- Therefore, the joint density is -->

<!-- \begin{align} -->

<!-- p(\textbf{x}|\lambda_1, \gamma_1,\lambda_2,\gamma_2, T) &= p(x_{c,1},\ldots,x_{c,n}|\lambda_2, \gamma_2) \times p(x_{t,1},\ldots,x_{t,m}|\lambda_1, \gamma_1,\lambda_2,\gamma_2, T) \\ -->

<!-- &= p(x_{c,1},\ldots,x_{c,n}|\lambda_2, \gamma_2) \times p(x_{t,1},\ldots,x_{t,j}|\lambda_2, \gamma_2) \times p(x_{t,j+1},\ldots,x_{t,m}|\lambda_1, \gamma_1,\lambda_2,\gamma_2, T), -->

<!-- \end{align} -->

<!-- where $x_{t,j}$ is the largest observed $x$ in the treatment group before T. -->

<!-- The joint likelihood is  -->

<!-- \begin{align} -->

<!-- \mathcal{L}(\lambda_1, \gamma_1,\lambda_2,\gamma_2, T|x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m}) &= \prod^n_{i=1} \frac{\gamma_2(\lambda_2x_{c,i})^{\gamma_2}\text{exp}\{-(\lambda_2x_{c,i})^{\gamma_2}\}}{x_{c,i}} \times \prod^j_{k=1}  \frac{\gamma_2(\lambda_2x_{t,k})^{\gamma_2}\text{exp}\{-(\lambda_2x_{t,k})^{\gamma_2}\}}{x_{t,k}} \times \prod^m_{p=j+1} -->

<!-- \gamma_1\lambda_1^{\gamma_1}x_{t,p}^{\gamma_1-1}\text{exp}\{-\lambda_1^{\gamma_1}(x_{t,p}^{\gamma_1}-T^{\gamma_1})-(T\lambda_2)^{\gamma_2}\} \\ -->

<!-- &= \gamma_2^n\lambda_2^{n\gamma_2}\{\prod_{i=1}^nx_{c,i}\}^{\gamma_2}\text{exp}\{-\lambda_2^{\gamma_2}\sum^n_{i=1}(x_{c,i}^{\gamma_2})\}\prod^n_{i=1}\frac{1}{x_{c,i}} \times \gamma_2^j\lambda_2^{j\gamma_2}\{\prod_{k=1}^jx_{t,k}\}^{\gamma_2}\text{exp}\{-\lambda_2^{\gamma_2}\sum^j_{k=1}(x_{t,k}^{\gamma_2})\}\prod^j_{k=1}\frac{1}{x_{t,k}} \times \gamma_1^{m-j}\lambda_1^{\gamma_1(m-j)}\{\prod^m_{p=j+1}x_{t,p}\}^{\gamma_1-1}\text{exp}\{-\lambda_1^{\gamma_1}\sum^m_{p=j+1}(x_{t,p}^{\gamma_1})+(m-j)\big[(\lambda_1T)^{\gamma_1}-(\lambda_2T)^{\gamma_2}\big]\} -->

<!-- \end{align} -->

<!-- We find the log-likelihood -->

<!-- \begin{align} -->

<!-- \text{log} \mathcal{L}(\lambda_1, \gamma_1,\lambda_2,\gamma_2, T|x_{c,1},\ldots,x_{c,n}, x_{t,1}, \ldots, x_{t, m}) &= n\text{log}\gamma_2 + n\gamma_2\text{log}\lambda_2+\gamma_2\sum^n_{i=1}\text{log}x_{c,i}-\lambda_2^{\gamma_2}\sum^{n}_{i=1}x_{c,i}^{\gamma_2}+\sum^{n}_{i=1}\text{log}\frac{1}{x_{c,i}} + j\text{log}\gamma_2 + j\gamma_2\text{log}\lambda_2+\gamma_2\sum^j_{k=1}\text{log}x_{t,k}-\lambda_2^{\gamma_2}\sum^{j}_{k=1}x_{t,k}^{\gamma_2}+\sum^{j}_{k=1}\text{log}\frac{1}{x_{t,k}}+(m-j)\text{log}\gamma_1+\gamma_1(m-j)\text{log}\lambda_1+(\gamma_1-1)\sum^m_{p=j+1}\text{log}x_{t,p} - \lambda_1^{\gamma_1}\sum^m_{p=j+1}(x_{t,p}^{\gamma_1})+(m-j)\big[(\lambda_1T)^{\gamma_1}-(\lambda_2T)^{\gamma_2}\big] -->

<!-- \end{align} -->

If we let $\gamma_1=\gamma_2$ the hazard ratio is

$$
\text{HR}=
\begin{cases}
 1,,& t\leq T \\
 (\frac{\lambda_1}{\lambda_2})^{\gamma_2}  ,& t>T
 \end{cases} 
$$

In consultation with experts, we have elicited the following distributions:

$$T \sim N(6, 1)$$ $$\text{post-delay HR} \sim Be(6.6, 4.5)$$

These distributions can be seen below

```{r, eval=T, echo=F}
x <- seq(0, 12, by=0.01)
y <- dnorm(x, 6, sd = 1)
plot(x, y, type="l", xlab="T", ylab="density")


x <- seq(0, 1, by=0.01)
y <- dbeta(x, 6.6, 4.5)
plot(x, y, type="l", xlab="post-delay HR", ylab="density")
```

We can simulate some data according to some underlying parameters (in practice, we would not know these), combine the data with our prior distributions to obtain posterior distributions for $T$ and post-delay HR.

This code chunk does this process once, for illustration. 

```{r}

HR <- rnorm(1, 0.7, sd = 0.1)
bigT <- rgamma(1, 65.5, 10.9)
gamma1 <- gamma2 <- 0.8
lambda2 <- 0.08
lambda1  <- lambda2*HR^(1/gamma2)
IATime <- 40

#Sample sizes in each group
n1 <- 150
n2 <- 150

#When is the final analysis time?
#trialLength <- 60

#Simulating the control and treatment data - again, we would not normally know the underlying structure, but for illustration this is okay
#Control
controldata <- rweibull(n1, gamma2, 1/lambda2)
#Treatment
CP <- exp(-(lambda2*bigT)^gamma2)[[1]]
u <- runif(n2)
suppressWarnings(treatmentdata <- ifelse(u>CP, (1/lambda2)*exp(1/gamma2*log(-log(u))), exp((1/gamma1)*log(1/(lambda1^gamma1)*(-log(u)-(lambda2*bigT)^gamma2+lambda1^gamma1*bigT*gamma1)))))

combinedData <- data.frame(time = c(controldata, treatmentdata), group = c(rep("Control", n1), rep("Treatment", n2)))

combinedData$event <- combinedData$time<IATime

combinedData$time[combinedData$time>IATime] <- IATime
# 
controlkm <- survfit(Surv(time, event)~group, data = combinedData)
# 
plot(controlkm, col=c("blue", "red"))

legend("topright", legend = c("Control", "Treatment"), col = c("blue", "red"), lty=1)

n <- n1
m <- n1+n2

#JAGS code which calculates posterior distributions

modelstring="

data {
  for (j in 1:m){
    zeros[j] <- 0
  }
}

model {
  C <- 10000
  for (i in 1:n){
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), -(lambda2*datTimes[i])^gamma2)
  }
  for (i in (n+1):m){                                                                                                             
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, ifelse(datTimes[i]<bigT, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), log(gamma2)+gamma2*log(lambda1)+(gamma2-1)*log(datTimes[i])-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)-(bigT*lambda2)^gamma2), 
      ifelse(datTimes[i]<bigT, -(lambda2*datTimes[i])^gamma2, -(lambda2*bigT)^gamma2-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)))
  }
  
    lambda2 ~ dbeta(1,1)T(0,)
    gamma2 ~ dbeta(1,1)T(0,)
    HR ~ dnorm(0.7, 10)T(0,)
    bigT ~ dgamma(65.5, 10.9)
    lambda1 <- lambda2*pow(HR, 1/gamma2)
    
    }
"

model = jags.model(textConnection(modelstring), data = list(datTimes = combinedData$time, datEvent = combinedData$event, n= n, m=m), quiet = T) 
                   
                   
update(model, n.iter=500)
output=coda.samples(model=model, variable.names=c("HR", "bigT"), n.iter = 5000)


summary(output)
plot(output)

```

This next code chunk repeats the above process 100 times (for each interim analysis time), it also varies the interim analysis time. 

```{r, eval=T, echo=T, cache=T}
#set.seed(53)
#Setting up the correct parameters - we would not know this in practice

TPPIA <- function(IATime){
  TPPval <- rep(NA, 101)
  HRval <- rep(NA, 101)
  TPPval[1] <- IATime
for (i in 2:101){
HR <- rnorm(1, 0.7, sd = 0.1414)
bigT <- rgamma(1, 65.5, 10.9)
gamma1 <- gamma2 <- 0.8
lambda2 <- 0.08
lambda1  <- lambda2*HR^(1/gamma2)

HRval[i] <- HR
#Sample sizes in each group
n1 <- 400
n2 <- 400

#When is the final analysis time?
#trialLength <- 60

#Simulating the control and treatment data - again, we would not normally know the underlying structure, but for illustration this is okay
#Control
controldata <- rweibull(n1, gamma2, 1/lambda2)
#Treatment
CP <- exp(-(lambda2*bigT)^gamma2)[[1]]
u <- runif(n2)
suppressWarnings(treatmentdata <- ifelse(u>CP, (1/lambda2)*exp(1/gamma2*log(-log(u))), exp((1/gamma1)*log(1/(lambda1^gamma1)*(-log(u)-(lambda2*bigT)^gamma2+lambda1^gamma1*bigT*gamma1)))))

combinedData <- data.frame(time = c(controldata, treatmentdata), group = c(rep("Control", n1), rep("Treatment", n2)))

combinedData$event <- combinedData$time<IATime

combinedData$time[combinedData$time>IATime] <- IATime
# 
# controlkm <- survfit(Surv(time, event)~group, data = combinedData)
# 
# plot(controlkm, col=c("blue", "red"))

#legend("topright", legend = c("Control", "Treatment"), col = c("blue", "red"), lty=1)

n <- n1
m <- n1+n2

#JAGS code which calculates posterior distributions

modelstring="

data {
  for (j in 1:m){
    zeros[j] <- 0
  }
}

model {
  C <- 10000
  for (i in 1:n){
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), -(lambda2*datTimes[i])^gamma2)
  }
  for (i in (n+1):m){                                                                                                             
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, ifelse(datTimes[i]<bigT, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), log(gamma2)+gamma2*log(lambda1)+(gamma2-1)*log(datTimes[i])-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)-(bigT*lambda2)^gamma2), 
      ifelse(datTimes[i]<bigT, -(lambda2*datTimes[i])^gamma2, -(lambda2*bigT)^gamma2-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)))
  }
  
    lambda2 ~ dbeta(1,1)T(0,)
    gamma2 ~ dbeta(1,1)T(0,)
    HR ~ dnorm(0.7, 50)T(0,)
    bigT ~ dgamma(65.5, 10.9)
    lambda1 <- lambda2*pow(HR, 1/gamma2)
    
    }
"

model = jags.model(textConnection(modelstring), data = list(datTimes = combinedData$time, datEvent = combinedData$event, n= n, m=m), quiet = T) 
                   
                   
update(model, n.iter=500)
output=coda.samples(model=model, variable.names=c("HR", "bigT"), n.iter = 2000)

# 
#summary(output)
#plot(output)
print(IATime)
print(i)
TPPval[i] <- mean(output[[1]][,1]<0.65)
}


return(list(TPPval = TPPval, HRval = HRval))

}

IATimeVec <- seq(5, 70, by=10)
#IATimeVec <- 10

outputvec <- sapply(X = IATimeVec,FUN= TPPIA)

#plot(IATimeVec, outputvec)

#We now need to sample from these posterior distributions
#We sample future patient data

# BPPvec <- rep(NA, 100)
# 
# for (i in 1:100){
#   #Calculating how many control events are remaining
# 
# noremainingcontrolevents <-  combinedData %>%
# filter(group=="Control"& event==F) %>%
#   count() %>%
#    pull()
# 
# #What is the  survival probability at time = IAtime?
#  
# controlIAsurv <- exp(-(lambda2*IATime)^gamma2)
# 
# u <- runif(noremainingcontrolevents, 0, controlIAsurv)
# 
# #Simulating the remaining control data
# 
# remainingcontrolevents <- data.frame(time = (1/lambda2)*(-log(u))^(1/gamma2), group = rep("Control", noremainingcontrolevents), event = rep(F, noremainingcontrolevents))
# 
# 
# #Now need to simulate the remaining treatment events
# #Using the posteriors for T and HR
# 
# noremainingtreatmentevents <- combinedData %>%
# filter(group=="Treatment"& event==F) %>%
#   count() %>%
#    pull()
# 
# 
# sampledHR <- sample(output[[1]][,1], size =1)
# sampledT <- sample(output[[1]][,2], size =1)
# 
# # print(sampledHR)
# # print(sampledT)
# 
# sampledlambda1 <- lambda2*sampledHR^(1/gamma2)
# 
# #What is the  survival probability at time = IAtime?
# treatmentIAsurv <- exp(-(lambda2*sampledT)^gamma2-sampledlambda1^gamma1*(IATime^gamma1-sampledT^gamma1))
# 
# u <- runif(noremainingtreatmentevents, 0, treatmentIAsurv)
# 
# #Sampling the remaining treatment events
# remainingtreatmentevents <- data.frame(time = ((1/(lambda1^gamma1))*((lambda1*sampledT)^gamma1-log(u)-(lambda2*sampledT)^gamma2))^(1/gamma1), group = rep("Treatment", noremainingtreatmentevents), event = rep(F, noremainingtreatmentevents))
# 
# #Now we need to combine the new simulated data with the original data set
# 
# original <- combinedData[combinedData$event==T,]
# 
# finalDataSet <- rbind(original, remainingcontrolevents, remainingtreatmentevents)
# 
# finalDataSet$event <- finalDataSet$time<trialLength
# 
# finalDataSet$time[finalDataSet$time>IATime] <- trialLength
# 
# test <- survdiff(Surv(time, event)~group, data = finalDataSet)
# 
# BPPvec[i] <- test$chisq > qchisq(0.95, 1)
# }
# 
# mean(BPPvec)


```

There is a question of what to do at this interim analysis stage, we could use these posteriors - especially the one for HR - to look at how much more informed we are at this interim analysis stage than we were compared to before the trial. We could also sample future observations from these posterior distributions and then combine these simulated future observations with the observations seen at the interim analysis stage to calculate Bayesian predictive probabilities.

```{r}

visualisingMCMCOutput <- function(data_vec, changing){
  len_dat <- length(data_vec)/2
  print(len_dat)
  len_dat_1 <- length(outputvec[,1]$TPPval)
  print(len_dat_1)
  if (len_dat%%2==0){
    par(mfrow=c(2,len_dat/2))
  } else {
    par(mfrow=c(2,(len_dat+1)/2))
  }
  densityVec <- rep(NA, length = length(len_dat))
  for (i in 1:len_dat){
     h <- hist(data_vec[,i]$TPPval[2:len_dat_1], xlim=c(0,1), freq=F)
     densityVec[i] <- max(h$density)
  }
  for (i in 1:len_dat){
    h <- hist(data_vec[,i]$TPPval[2:len_dat_1], xlim=c(0,1), ylim=c(0, max(densityVec)),
              main=paste0("Histogram, ", changing, "= ", data_vec[,i]$TPPval[1]), xlab = "Proportion of posterior < TPP", freq=F)
    #print(mean(data_vec[,i]$TPPval[2:len_dat_1]))
  }
  
  # if (len_dat%%2==0){
  #   par(mfrow=c(2,len_dat/2))
  # } else {
  #   par(mfrow=c(2,(len_dat+1)/2))
  # }
  # for (j in 1:len_dat){
  #   plot(data_vec[,j]$HRval[2:len_dat_1], data_vec[,j]$TPPval[2:len_dat_1], ylab = "Proportion of posterior < TPP",
  #        xlab = "Sampled HR", xlim = c(0,1), main = paste0("Plot, t = ", data_vec[,j]$TPPval[1]))
  # }
}



```

This next code chunk uses an approximation of the hazard ratio - instead of doing "full" MCMC on it.
This changes the interim analysis time.

```{r}

TPPIA <- function(IATime){
  n <- 10001
  TPPval <- rep(NA, n)
  HRval <- rep(NA, n)
  TPPval[1] <- IATime
  for (i in 2:n){
HR <- rnorm(1, 0.7, sd = 0.1)
bigT <- rgamma(1, 65.5, 10.9)
gamma1 <- gamma2 <- 0.8
lambda2 <- 0.08
lambda1  <- lambda2*HR^(1/gamma2)

#Sample sizes in each group
n1 <- 1000
n2 <- 1000

#What is the survival probability at t = IATime?
survprobcontrol <- exp(-(lambda2*IATime)^gamma2)
survprobtreatment <- exp(-(lambda2*bigT)^gamma2-lambda1^gamma1*(IATime^gamma1-bigT^gamma1))

#Expected number of events at t = 40
e1 <- (1-survprobcontrol)*n1
e2 <- (1-survprobtreatment)*n2

#print(e1)
#print(e2)
#print(sqrt(1/e1+1/e2))

HRval[i] <- HR
TPPval[i] <- pnorm(0.65, mean = HR, sd = sqrt(1/e1+1/e2), lower.tail = T)
}

return(list(TPPval = TPPval, HRval = HRval))

}

IATimeVec <- seq(1, 200 ,by=20)
#IATimeVec <- 10

outputvec <- sapply(X = IATimeVec,FUN= TPPIA)


visualisingMCMCOutput(outputvec, "t")


```

This next code chunk does exactly the same thing, but we keep interim analysis time constant and only change n. 

```{r}

TPPIA <- function(n1){
  n <- 10001
  IATime <- 40
  TPPval <- rep(NA, n)
  HRval <- rep(NA, n)
  TPPval[1] <- n1
  for (i in 2:n){
HR <- rnorm(1, 0.7, sd = 0.1)
bigT <- rgamma(1, 65.5, 10.9)
gamma1 <- gamma2 <- 0.8
lambda2 <- 0.08
lambda1  <- lambda2*HR^(1/gamma2)


#What is the survival probability at t = IATime?
survprobcontrol <- exp(-(lambda2*IATime)^gamma2)
survprobtreatment <- exp(-(lambda2*bigT)^gamma2-lambda1^gamma1*(IATime^gamma1-bigT^gamma1))

#Expected number of events at t = 40
e1 <- (1-survprobcontrol)*n1
e2 <- (1-survprobtreatment)*n1

#print(e1)
#print(e2)
#print(sqrt(1/e1+1/e2))

HRval[i] <- HR
TPPval[i] <- pnorm(0.65, mean = HR, sd = sqrt(1/e1+1/e2), lower.tail = T)
}

return(list(TPPval = TPPval, HRval = HRval))

}

nVec <- round(10^seq(0, 4, by=0.5))
#IATimeVec <- 10

outputvec <- sapply(X = nVec,FUN= TPPIA)

visualisingMCMCOutput(outputvec, "n")


```


```{r}

TPPIA <- function(n1, IATime){
  n <- 10001
  TPPval <- rep(NA, n)
  HRval <- rep(NA, n)
  TPPval[1] <- n1
  for (i in 2:n){
HR <- rnorm(1, 0.7, sd = 0.1)
bigT <- rgamma(1, 65.5, 10.9)
gamma1 <- gamma2 <- 0.8
lambda2 <- 0.08
lambda1  <- lambda2*HR^(1/gamma2)


#What is the survival probability at t = IATime?
survprobcontrol <- exp(-(lambda2*IATime)^gamma2)
survprobtreatment <- exp(-(lambda2*bigT)^gamma2-lambda1^gamma1*(IATime^gamma1-bigT^gamma1))

#Expected number of events at t = 40
e1 <- (1-survprobcontrol)*n1
e2 <- (1-survprobtreatment)*n1

#print(e1)
#print(e2)
#print(sqrt(1/e1+1/e2))

HRval[i] <- HR
TPPval[i] <- pnorm(0.65, mean = HR, sd = sqrt(1/e1+1/e2), lower.tail = T)
}

return(TPPval = TPPval)

}

nVec <- 3*(10^seq(0,3, by=0.5))
IATimeVec <- seq(5, 60 ,length = 6)

posteriorMatrix <-matrix(rep(list(), length(nVec)+length(IATimeVec)),nrow = length(nVec), ncol = length(IATimeVec))

for (i in 1:length(nVec)){
  for (j in 1:length(IATimeVec)){
    posteriorMatrix[i,j] <- list(TPPIA(nVec[i], IATimeVec[j]))
  }
}

```

par(mfrow=c(length(nVec),length(IATimeVec)))
densityMat <- matrix(NA, ncol = length(IATimeVec), nrow = length(nVec))
for (i in 1:length(nVec)){
  for (j in 1:length(IATimeVec)){
    h <- hist(unlist(posteriorMatrix[i,j])[2:1001], xlim=c(0,1), freq = F)
    densityMat[i,j] <- max(h$density)
  }
}

for (i in 1:length(nVec)){
  for (j in 1:length(IATimeVec)){
    hist(unlist(posteriorMatrix[i,j])[2:1001], xlim=c(0,1), xlab = "Proportion of posterior < TPP", ylim=c(0, max(densityMat)), main = paste0("Histogram: t = ", round(IATimeVec[j]), ", n = ", round(nVec[i])), freq = F)
  }
}

```{r}
posteriorMatrix <- readRDS("~/R Projects/PhD-Materials/posteriorMatrixJAGS.rds")
nVec <- 3*(10^seq(0,3))
IATimeVec <- seq(5, 60 ,length = 4)

par(mfrow=c(length(nVec),length(IATimeVec)))
densityMat <- matrix(NA, ncol = length(IATimeVec), nrow = length(nVec))
for (i in 1:length(nVec)){
  for (j in 1:length(IATimeVec)){
    h <- hist(unlist(posteriorMatrix[i,j])[2:1001], xlim=c(0,1), freq = F)
    densityMat[i,j] <- max(h$density)
  }
}

for (i in 1:length(nVec)){
  for (j in 1:length(IATimeVec)){
    hist(unlist(posteriorMatrix[i,j])[2:1001], xlim=c(0,1), xlab = "Proportion of posterior < TPP", ylim=c(0, max(densityMat)), main = paste0("Histogram: t = ", round(IATimeVec[j]), ", n = ", round(nVec[i])), freq = F)
  }
}

extremeMassMatrix <-matrix(rep(NA, length(nVec)+length(IATimeVec)),nrow = length(nVec), ncol = length(IATimeVec))
colnames(extremeMassMatrix) <- IATimeVec
rownames(extremeMassMatrix) <- nVec

for (i in 1:length(nVec)){
  for (j in 1:length(IATimeVec)){
   x <-  unlist(posteriorMatrix[i,j])
   extremeMassMatrix[i,j] <- mean(x>0.9|x<0.1)
  }
}


heatmap(extremeMassMatrix, Colv=NA, Rowv=NA, xlab = "interim time", scale = "none", ylab = "sample size", col= colorRampPalette(brewer.pal(8, "Oranges"))(25))


# legend(x="bottomright", legend = c("Min", "Med", "Large"),
#      fill=colorRampPalette(brewer.pal(8, "Oranges"))(3))
```



```{r}
#Instead of simulating the data, we can look at the data sets found from real clinical trials


JAGSFunc <- function(IATime){
  #Do the Borghaei data set first
controldata <- read.csv(file = paste0("~/R Projects/PhD-Materials/DTEPaper/data/Borghaei/IPD-control.csv"))
treatmentdata <- read.csv(file = paste0("~/R Projects/PhD-Materials/DTEPaper/data/Borghaei/IPD-treatment.csv"))
combinedData <- data.frame(time = c(controldata$Survival.time, treatmentdata$Survival.time), 
                             status = c(controldata$Status, treatmentdata$Status), 
                             group = c(rep("Control", nrow(controldata)), rep("Treatment", nrow(treatmentdata))))

#We can change the IA time here
#IATime <- 40
combinedData[combinedData$time>IATime,]$time <- IATime
combinedData[combinedData$time==IATime,]$status <- 0

#Plotting this data  
# kmfit <- survfit(Surv(time, status)~group, data = combinedData)
# plot(kmfit, conf.int = F, col=c("blue", "red"), xlab = "Time (months)", ylab = "Overall survival (%)", yaxt = "n")
# axis(2, at=seq(0, 1, by=0.1), labels=seq(0, 100, by=10))

#We can combine this data with vague priors to see whether we can capture the behaviour properly

n <- nrow(controldata)
m <- nrow(combinedData)

#JAGS code which calculates posterior distributions

modelstring="

data {
  for (j in 1:m){
    zeros[j] <- 0
  }
}

model {
  C <- 10000
  for (i in 1:n){
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), -(lambda2*datTimes[i])^gamma2)
  }
  for (i in (n+1):m){                                                                                                             
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, ifelse(datTimes[i]<bigT, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), log(gamma2)+gamma2*log(lambda1)+(gamma2-1)*log(datTimes[i])-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)-(bigT*lambda2)^gamma2), 
      ifelse(datTimes[i]<bigT, -(lambda2*datTimes[i])^gamma2, -(lambda2*bigT)^gamma2-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)))
  }
  
    lambda2 ~ dbeta(1,1)T(0,)
    gamma2 ~ dbeta(1,1)T(0,)
    HR ~ dnorm(0.7, 10)T(0,)
    bigT ~ dnorm(6, 1)T(0,)
    lambda1 <- lambda2*pow(HR, 1/gamma2)
    
    }
"

model = jags.model(textConnection(modelstring), data = list(datTimes = combinedData$time, datEvent = combinedData$status, n= n, m=m), quiet = T) 
                   
                   
update(model, n.iter=500)
output=coda.samples(model=model, variable.names=c("HR"), n.iter = 5000)

#summary(output)
#plot(output)

output <- as.numeric(output[[1]])
hist(output, freq = F, xlab = "HR", xlim=c(0.2, 1), ylim=c(0, 8), main = paste0("T = ", IATime))
}

par(mfrow=c(3, 4))

IATimeVec <- seq(5, 60, by=5)
sapply(IATimeVec, JAGSFunc)


```


```{r}
#Now try the Brahmer data set
JAGSFunc <- function(IATime){
controldata <- read.csv(file = paste0("~/R Projects/PhD-Materials/DTEPaper/data/Brahmer/IPD-control.csv"))
treatmentdata <- read.csv(file = paste0("~/R Projects/PhD-Materials/DTEPaper/data/Brahmer/IPD-treatment.csv"))
combinedData <- data.frame(time = c(controldata$Survival.time, treatmentdata$Survival.time), 
                             status = c(controldata$Status, treatmentdata$Status), 
                             group = c(rep("Control", nrow(controldata)), rep("Treatment", nrow(treatmentdata))))

#We can change the IA time here
#IATime <- 40
combinedData[combinedData$time>IATime,]$time <- IATime
combinedData[combinedData$time==IATime,]$status <- 0
  
# kmfit <- survfit(Surv(time, status)~group, data = combinedData)
# plot(kmfit, conf.int = F, col=c("blue", "red"), xlab = "Time (months)", ylab = "Overall survival (%)", yaxt = "n")
# axis(2, at=seq(0, 1, by=0.1), labels=seq(0, 100, by=10))

#We can combine this data with vague priors to see whether we can capture the behaviour properly

n <- nrow(controldata)
m <- nrow(combinedData)

#JAGS code which calculates posterior distributions

modelstring="

data {
  for (j in 1:m){
    zeros[j] <- 0
  }
}

model {
  C <- 10000
  for (i in 1:n){
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), -(lambda2*datTimes[i])^gamma2)
  }
  for (i in (n+1):m){                                                                                                             
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, ifelse(datTimes[i]<bigT, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), log(gamma2)+gamma2*log(lambda1)+(gamma2-1)*log(datTimes[i])-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)-(bigT*lambda2)^gamma2), 
      ifelse(datTimes[i]<bigT, -(lambda2*datTimes[i])^gamma2, -(lambda2*bigT)^gamma2-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)))
  }
  
    lambda2 ~ dbeta(1,1)T(0,)
    gamma2 ~ dbeta(1,1)T(0,)
    HR ~ dnorm(0.7, 10)T(0,)
    bigT ~ dnorm(3, 1)T(0,)
    lambda1 <- lambda2*pow(HR, 1/gamma2)
    
    }
"

model = jags.model(textConnection(modelstring), data = list(datTimes = combinedData$time, datEvent = combinedData$status, n= n, m=m), quiet = T) 
                   
                   
update(model, n.iter=1000)
output=coda.samples(model=model, variable.names=c("HR"), n.iter = 5000)

output <- as.numeric(output[[1]])
hist(output, freq = F, xlab = "HR", xlim=c(0.2, 1), ylim=c(0, 8), main = paste0("T = ", IATime))
}

par(mfrow=c(2, 4))

IATimeVec <- seq(5, 20, by=2)
sapply(IATimeVec, JAGSFunc)

```


```{r}
#Now try the Brahmer data set
controldata <- read.csv(file = paste0("~/R Projects/PhD-Materials/DTEPaper/data/Yen/IPD-control.csv"))
treatmentdata <- read.csv(file = paste0("~/R Projects/PhD-Materials/DTEPaper/data/Yen/IPD-treatment.csv"))
combinedData <- data.frame(time = c(controldata$Survival.time, treatmentdata$Survival.time), 
                             status = c(controldata$Status, treatmentdata$Status), 
                             group = c(rep("Control", nrow(controldata)), rep("Treatment", nrow(treatmentdata))))
  
kmfit <- survfit(Surv(time, status)~group, data = combinedData)
plot(kmfit, conf.int = F, col=c("blue", "red"), xlab = "Time (months)", ylab = "Overall survival (%)", yaxt = "n")
axis(2, at=seq(0, 1, by=0.1), labels=seq(0, 100, by=10))

#We can combine this data with vague priors to see whether we can capture the behaviour properly

n <- nrow(controldata)
m <- nrow(combinedData)

#JAGS code which calculates posterior distributions

modelstring="

data {
  for (j in 1:m){
    zeros[j] <- 0
  }
}

model {
  C <- 10000
  for (i in 1:n){
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), -(lambda2*datTimes[i])^gamma2)
  }
  for (i in (n+1):m){                                                                                                             
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <-  -l[i] + C
    l[i] <- ifelse(datEvent[i]==1, ifelse(datTimes[i]<bigT, log(gamma2)+gamma2*log(lambda2*datTimes[i])-(lambda2*datTimes[i])^gamma2-log(datTimes[i]), log(gamma2)+gamma2*log(lambda1)+(gamma2-1)*log(datTimes[i])-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)-(bigT*lambda2)^gamma2), 
      ifelse(datTimes[i]<bigT, -(lambda2*datTimes[i])^gamma2, -(lambda2*bigT)^gamma2-lambda1^gamma2*(datTimes[i]^gamma2-bigT^gamma2)))
  }
  
    lambda2 ~ dbeta(1,1)T(0,)
    gamma2 ~ dbeta(1,1)T(0,)
    HR ~ dnorm(0.7, 10)T(0,)
    bigT ~ dnorm(4, 1)T(0,)
    lambda1 <- lambda2*pow(HR, 1/gamma2)
    
    }
"

model = jags.model(textConnection(modelstring), data = list(datTimes = combinedData$time, datEvent = combinedData$status, n= n, m=m), quiet = T) 
                   
                   
update(model, n.iter=1000)
output=coda.samples(model=model, variable.names=c("HR", "bigT"), n.iter = 5000)

summary(output)
plot(output)

```



